Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-19 latex/article.tex

Representing a hierachical structure on GPU is usually challenging
since the GPU memory is limited and non consistent \todo{fix coherent/aligned/coalesced?}
memory access (\emph{e.g.} tree traversal using pointer dereferencing)
may lead to huge payload from cache misses.  Efficient spatial tree
encoding can be achieved using pointerless strictures. We use here
linear quadtrees or octrees to represent classical quadtree/octreee
trees, see for instance Gargantini
\cite{gargantini1982effective}. This structure consists in indexing
each cell by a \emph{Morton code}: the code of children cells are
defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). For a given cell, its
code encodes its position with respect to its parent cell and its
complete path to the tree root. Hence, whatever their depth, all cell
codes are simply stored in a linear vector on memory. Furthermore,
cell operations such as subdivision, merging, neighbors access can be
efficiently implemented using bitwise operations on the Morton code.
In the following, we use the GPU implementation proposed by Dupuy
\textit{et al.}  \cite{dupuy2014quadtrees}. In such approach, Morton
codes and bitwise operations are optimized to match with GPU hardware
requirements.

@@

As illustrated in Figure
\ref{fig_quadtree_partitionning}-\emph{middle}, we propose a viewpoint
dependent criterion to decide if a cell needs to be refined: the
closer we are to the camera, the finer the cells are. In addition to
the distance criterion, we define an angular criterion to refine cells
within the camera frustrum. When rendering a mesh, a key stage is the
rasterization one which digitizes triangles projected onto the
screen. As a consequence, having several triangles whose projections
lie inside the same pixel is something we should prevent. Hence,
distance criterion can be set in order to both control the geometrical
approximation and to ensure that cells project onto more than one
pixel. This does not prevent all triangles within such cells to be
projected into more than one pixel but gives us a fast criterion to
decide if a cell should be split or not.  Figure \ref{fig_lod_octree}
illustrates the level of details (LoD for short criterion) we
use. More precisely, in dimension 2, if $\alpha$ denotes the viewing
angle, an object at distance $d$ from the camera with diameter greater
than $2\cdot d\cdot\tan(\alpha)$ is projected into more than one pixel
(see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the distance
criterion is based on the ratio (\emph{visibility ratio} in the
following)between the cell diameter $l(c)$ (power of 2 depending on
the depth), and $2\cdot d_c\cdot\tan(\alpha)$, $d_c$ being the
distance between the cell center and the camera. For a given cell $c$,
split and merge decision are based on this visibility ratio:
\begin{itemize}
\item $c$ is split if its children cells have a visibility ratio
  greater than  constant $k$\,;
\item $c$ and its sibling cells are merged if their parent cell $c'$
  have av visibility ratio lower than $k$\;
\item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=4cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=4cm]{viewlod2_small}~~~~~
    \includegraphics[width=3.5cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and
    angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is
  updated between each frame \emph{(right)}.}
  \label{fig_lod_octree} 
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please not that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.

>>>>

Representing a hierachical structure on GPU is usually challenging
since the GPU memory is limited and non consistent \todo{fix coherent/aligned/coalesced?}
memory access (\emph{e.g.} tree traversal using pointer dereferencing)
may lead to huge payload from cache misses.  Efficient spatial tree
encoding can be achieved using pointerless structures. We use
linear quadtrees or octrees to represent classical quadtree/octreee
trees, see for instance Gargantini
\cite{gargantini1982effective}. This structure consists in indexing
each cell by a \emph{Morton code}: the code of children cells are
defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes
 its position with respect to its parent cell and its
complete path to the tree root. Hence, whatever their depth, all cell
codes are simply stored in a linear vector on memory. Furthermore,
cell operations such as subdivision, merging, neighbors access can be
efficiently implemented using bitwise operations on the Morton code.
In the following, we use the GPU implementation proposed by Dupuy
\textit{et al.}  \cite{dupuy2014quadtrees}. By using their approach, Morton
codes and bitwise operations are optimized to match with GPU hardware
requirements.

@@

As illustrated in Figure
\ref{fig_quadtree_partitionning}-\emph{middle}, we propose a viewpoint
dependent criterion to decide if a cell needs to be refined: the
closer we are to the camera, the finer the cells are. In addition to
the distance criterion, we define an angular criterion to refine cells
within the camera frustrum. When rendering a mesh, a key stage is the
rasterization which digitizes triangles projected onto the
screen. As a consequence, having several triangles whose projections
lie inside the same pixel is something we should prevent. Hence,
distance criterion can be set in order to both control the geometrical
approximation and to ensure that cells project onto more than one
pixel. This does not prevent all triangles within such cells to be
projected into more than one pixel but gives us a fast criterion to
decide if a cell should be split or not.  Figure \ref{fig_lod_octree}
illustrates the level of details (LoD for short criterion) we
use. More precisely, in dimension 2, if $\alpha$ denotes the viewing
angle, an object at distance $d$ from the camera with diameter greater than
\todo{Extract formulas from text with dots list} $2\cdot d\cdot\tan(\alpha)$
 is projected into more than one pixel
(see Figure~\ref{fig_lod_octree}-\emph{left}). Hence, the distance
criterion is based on the ratio (\emph{visibility ratio} in the
following)between the cell diameter $l(c)$ (power of 2 depending on
the depth), and 
$2\cdot d_c\cdot\tan(\alpha)$ \todo{idem}
,$d_c$ being the distance between the cell center and the camera. For a given cell $c$,
split and merge decision are based on this visibility ratio:
\begin{itemize}
\item $c$ is split if its children cells have a visibility ratio
  greater than  constant $k$\,;
\item $c$ and its sibling cells are merged if their parent cell $c'$
  have av visibility ratio lower than $k$\;
\item otherwise, the cell $c$ stays for the next frame.
\end{itemize}
%
\begin{figure}[!h]
  \begin{center}
    \begin{overpic}[width=4cm]{figs/criterion}
      \put(33,40){$\alpha$}
      \put(50,45){$d_c$}
      \put(95,65){$l(c)$}
    \end{overpic}~~~~~
    \includegraphics[width=4cm]{viewlod2_small}~~~~~
    \includegraphics[width=3.5cm]{subdivision}
  \end{center}
  \caption{Notations \emph{(left)} and adaptive meshing in dimension 2 using the LoD distance and
    angular criterion \emph{(middle)}. Based on LoD criterion
    evaluation, the cell buffer is
  updated between each frame \emph{(right)}.}
  \label{fig_lod_octree} 
\end{figure}
%
In addition to the distance criterion, we add an angular term in order
to prevent subdivision of cells far from the camera frustrum (see
Figure~\ref{fig_lod_octree}-\emph{middle}). We do not go into the
details but please note that stay, split and merge decisions of all cells are computed in
parallel from all cell morton codes.
